---
editor:
  markdown:
    wrap: 72
---

# 3. Sem 3 Peer Mentors {.unnumbered}

```{r}
#| include: false
source("_setup.R")
```

*Who is in this cohort, and what does my stream do?*

## You are the mentor now {.unnumbered}

You are now a peer mentor for the FRI Semester 3 course, and the thing you are
developing is no longer your own research skill but your ability to grow other
people's. Those are genuinely different skills. Being good at research does not
automatically make someone good at mentoring, and the gap between the two is
where this guide lives.

The measures in the chapters ahead are not tests. Nobody is scored as a better
or worse mentor. Each one names a style or an orientation you already have,
shows you where it sits relative to the rest of your cohort, and asks you to
decide whether it is serving your mentees this week.

## This cohort {.unnumbered}

There are **25 peer mentors** in the Semester 3 cohort, across three streams.

```{r}
#| label: ch03-cohort-table
#| echo: false
dat |>
  count(Stream = STREAM, name = "Peer mentors") |>
  kbl(caption = "Semester 3 peer mentors by stream") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)
```

## What this cohort wants {.unnumbered}

The first question on the survey asked how important it is to you to become a
**more effective peer mentor** by the end of this semester. This is the one
score in the guide where the answer is not a style but a commitment, and it is
worth seeing what your cohort said.

```{r}
#| label: ch03-effective-plot
#| echo: false
#| fig-width: 7
#| fig-height: 4
# A five-point single item, so this is a bar chart over the response options -
# not norm_facets(), whose histogram bins would straddle the response values.
eff <- tibble::tibble(
  Importance = factor(1:5, levels = 1:5,
    labels = c("Not\nimportant", "Slightly", "Moderately", "Important",
               "Very\nimportant")),
  n = as.integer(table(factor(dat$EFFECTIVE, levels = 1:5))))

ggplot(eff, aes(Importance, n)) +
  geom_col(fill = green, width = 0.62) +
  geom_text(aes(label = n), vjust = -0.55, size = 4, color = neutral_dark) +
  scale_y_continuous(breaks = seq(0, 14, 2), expand = expansion(mult = c(0, 0.12))) +
  labs(x = "How important is becoming a more effective peer mentor?",
       y = "Peer mentors") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title.x = element_text(color = green, face = "bold"))
```

Almost everyone rated it **Important** or **Very important**. That is a real
starting condition: the room already wants this.

::: callout-note
## Why this one has no bands

Every other chapter shades the distribution into Low / Mid / High thirds. This
one does not. Read this figure as a group portrait, not as a personal score.
:::

## Your stream {.unnumbered}

The mindsets in this guide are the same for everyone. Good mentoring looks much
the same whether your team is running assays, analyzing health data, or building
visualizations — the chapters ahead apply to all three streams without
modification.

**What differs is how your stream uses the guide.** Streams set their own pace
through the chapters, hold their peer mentor meetings at different times, and
build the weekly reflection into their course in different ways. Your research
educator decides which chapter you are working on in a given week, how the
mindset is discussed when you meet, and where your journal entries go.

Your research educator will tell you which chapter to read each week, when your
peer mentor meetings happen, and how to submit your journal entries.

**Ask your research educator for details.**

## Reflect and Plan

::: {.callout-warning}
## Journal
- **Purpose:** What made a peer mentor helpful to you when you were the
  first-year student? Name something specific they did.
- **Goal:** You rated how important it is to become a more effective mentor.
  What would "more effective" actually look like for you by December?
:::
