# 6. Leadership Focus {.unnumbered}

```{r}
#| include: false
source("_setup.R")
```

*Is my leadership style to promote or prevent for the team?*

## The Mindset(s) {.unnumbered}

The same regulatory focus from Chapter 5, now applied to how you lead. Each pole covers the same
three leadership functions: goal framing, problem solving, and error management. (Feedback is
its own chapter: **Chapter 7, Feedback Orientation**.)

- [**Promotion Leadership**]{style="color:#5A71A8;"}: frames goals around gains, builds on
  what works, and encourages useful risks.
  - *Survey item:* "I plan to motivate my team by framing our goals around the gains and growth we can achieve."
- [**Prevention Leadership**]{style="color:#453A98;"}: frames goals around setbacks to
  avoid, fixes what is going wrong, and emphasizes accuracy.
  - *Survey item:* "I plan to motivate my team by framing our goals around the problems and setbacks we need to avoid."

**Polarity:** Promotion ↔ Prevention (team leadership).

**Measure:** Promotion/Prevention Team Leadership Scale (adapted from regulatory focus
theory). Six items rated 1–7, three per subscale.

## Find Your Score {.unnumbered}

Use your password to find your score(s).

::: {.callout-note}
## How your Preference is decided
Your **Preference** comes from comparing your two **percentiles**. If your Promotion percentile
sits more than 20 points above your Prevention percentile you are **Promotion-leaning**; more
than 20 points below, **Prevention-leaning**; within 20 points of each other, **Balanced**.

Unlike the point budget in Chapter 4, these two scales are **independent** — you are not
trading one off against the other, and you can be high or low on both. So *Balanced* here does
not mean lukewarm. It means you bring both framings to your team at a similar strength relative
to your peers, which is often the more useful place to lead from.
:::

```{r}
#| label: ch06-leadership-focus-compute
#| include: false
w3 <- dat |> transmute(Password = PASSWORD,
  `Promotion Leadership`  = rowMeans(across(c(Lead_Promo1, Lead_Promo2, Lead_Promo3))),
  `Prevention Leadership` = rowMeans(across(c(Lead_Prev1, Lead_Prev2, Lead_Prev3))))

# Preference compares the two PERCENTILES, using the same +/-20 rule as Chapter 4.
# Unlike the Chapter 4 budget, these two scales are independent (they do not share a
# fixed pool), so a mentor can be high or low on both - which is why the middle band
# here is wide: holding both focuses about equally is a real result, not a fudge.
# Rounded to the values printed in the table so the label always agrees with them.
pct_promo <- round(dplyr::percent_rank(w3$`Promotion Leadership`)  * 100)
pct_prev  <- round(dplyr::percent_rank(w3$`Prevention Leadership`) * 100)
pref3 <- factor(dplyr::case_when(
  pct_promo - pct_prev >  20 ~ "Promotion-leaning",
  pct_promo - pct_prev < -20 ~ "Prevention-leaning",
  TRUE                       ~ "Balanced"),
  levels = c("Prevention-leaning", "Balanced", "Promotion-leaning"))
```

```{r}
#| label: ch06-leadership-focus-table
#| echo: false
score_table(w3$Password, list(`Promotion Leadership`=w3$`Promotion Leadership`,
  `Prevention Leadership`=w3$`Prevention Leadership`),
  "6. Leadership Focus: your composite scores (1–7)",
  extra = list(Preference = pref3))
```

## Locate Your Score {.unnumbered}

Locate your score(s) relative to other peer mentors!

```{r}
#| label: ch06-leadership-focus-plot
#| echo: false
#| fig-width: 9
#| fig-height: 3.8
norm_facets(list(`Promotion Leadership`=w3$`Promotion Leadership`,
  `Prevention Leadership`=w3$`Prevention Leadership`), xlab="Score (1–7)",
  scale = c(1, 7))
```

## Reflect and Plan

::: {.callout-warning}
## Journal
- **Score:** How does reflecting on your score help you think about who you are as a mentor?
- **Behavior:** To be the best mentor you can be, what is one new thing you will try to do this week for mentees and/or your team?
:::
