# 9. Leadership Role {.unnumbered}

```{r}
#| include: false
source("_setup.R")
```

*What is my leadership role?*

## The Mindset(s) {.unnumbered}

Leading a team is not one job but many. Morgeson, DeRue & Karam (2010) describe team
leadership as a set of distinct **functions** a leader can take on. This week you decide which
of ten functions you see as **your role** as a peer mentor. They fall into two groups:

- **Task leadership**: getting the work done: planning, tracking, and training.
  - *Survey item:* "As a peer mentor, it is my role to help them plan."
- **Relational leadership**: supporting the people: monitoring, motivating, developing, and
  holding the team together.
  - *Survey item:* "As a peer mentor, it is my role to motivate the team to excel."

**Polarity:** Task ↔ Relational leadership.

**Measure:** Mentoring Leadership Functions inventory, adapted from Morgeson et al. (2010),
Table 2. The stem is *"As a peer mentor, it is my role to…"* and you rate each of ten
functions 1–10. You have a **budget of 50 points** to spread across the ten, so you have to
prioritize which functions are most central to your role. Your **Task** and **Relational**
scores are the total points you place on each group's functions (out of 50).

- **Task functions:** help them plan · keep track of assignment due dates · train the team on
  specific research skills · provide feedback to teammates · evaluate the performance of the team
- **Relational functions:** monitor the group chat for help · motivate the team to excel ·
  encourage team members to self-manage · resolve team conflict · build friendships

## Find Your Score {.unnumbered}

Use your password to find your score(s).

::: {.callout-note}
## How your Preference is decided
Your **Preference** comes from comparing your two **percentiles**. If your Task percentile sits
more than 20 points above your Relational percentile you are **Task-leaning**; more than 20
points below, **Relational-leaning**; within 20 points of each other, **Balanced**.

The percentiles are used rather than the raw point totals because the 50 points are a fixed
budget: judging against an even 25/25 split would say more about how the ten functions are
worded than about how you see your role. Comparing percentiles asks which way you lean
**relative to your peers**.
:::

```{r}
#| label: ch09-leadership-role-compute
#| include: false
# Task1-5 / Relational1-5 per the Semester 3 instrument labelling. Note that
# "evaluate the performance of the team" (LEADROLE_8) is Task5 here; it counted as
# a relational function in an earlier version of this instrument, so scores from an
# earlier administration are not comparable.
w6 <- dat |> transmute(Password = PASSWORD,
  Task       = LEADROLE_1 + LEADROLE_4 + LEADROLE_6 + LEADROLE_7 + LEADROLE_8,
  Relational = LEADROLE_5 + LEADROLE_9 + LEADROLE_10 + LEADROLE_11 + LEADROLE_12)

# Preference compares the two PERCENTILES, using the same +/-20 rule as Chapters 4 and 6.
# Like Chapter 4 this is a fixed budget, so the two percentiles mirror each other and
# the comparison amounts to which side of the cohort you sit on. Rounded to the values
# printed in the table so the label always agrees with them.
pct_task <- round(dplyr::percent_rank(w6$Task)       * 100)
pct_rel  <- round(dplyr::percent_rank(w6$Relational) * 100)
pref6 <- factor(dplyr::case_when(
  pct_task - pct_rel >  20 ~ "Task-leaning",
  pct_task - pct_rel < -20 ~ "Relational-leaning",
  TRUE                     ~ "Balanced"),
  levels = c("Relational-leaning", "Balanced", "Task-leaning"))
```

```{r}
#| label: ch09-leadership-role-table
#| echo: false
score_table(w6$Password, list(Task=w6$Task, Relational=w6$Relational),
  "9. Leadership Role: your point totals (out of 50)",
  extra = list(Preference = pref6))
```

## Locate Your Score {.unnumbered}

Locate your score(s) relative to other peer mentors!

```{r}
#| label: ch09-leadership-role-plot
#| echo: false
#| fig-width: 9
#| fig-height: 3.8
norm_facets(list(Task=w6$Task, Relational=w6$Relational),
  xlab="Points (out of 50)", scale = c(0, 50), by = 10)
```

::: {.callout-important}
## About this measure
The ten functions are drawn from Morgeson, DeRue & Karam's (2010) functional model of team
leadership, but the mentoring-specific wording, the importance framing, and the 50-point
budget are **adapted for this course**. Because the points are a fixed budget, spending more
on one group of functions necessarily means spending less on the other, and the two totals
always sum to exactly 50. Task and Relational cover five functions each, so the two scores are
directly comparable — but still read each one against the cohort distribution above rather than
against the scale. Because the two are complements, their percentiles always look like
opposites: a high Task percentile forces a low Relational one. Re-check reliability before
reporting results beyond this guide.
:::

## Reflect and Plan

::: {.callout-warning}
## Journal
- **Score:** How does reflecting on your score help you think about who you are as a mentor?
- **Behavior:** To be the best mentor you can be, what is one new thing you will try to do this week for mentees and/or your team?
:::
