# 10. Communication Style {.unnumbered}

```{r}
#| include: false
source("_setup.R")
```

*What is my communication style?*

## The Mindset(s) {.unnumbered}

Nonviolent Communication describes three behaviors that keep a hard conversation connected
rather than combative.

- [**Self-connection**]{style="color:#007F85;"}: recognizing that your emotions come from
  your own thoughts and unmet needs, and staying present enough under stress to notice a
  reaction without acting on it.
  - *Survey item:* "Knowing that my emotions stem from my own thoughts and unmet needs, my emotions are not caused by others."
- [**Authentic self-expression**]{style="color:#5A71A8;"}: stating your feelings, needs, and
  requests sincerely, without accusation, and staying open to any response.
  - *Survey item:* "Being responsible for my own experience, and willing to express my feelings and needs sincerely without accusation or criticism."
- [**Empathic listening**]{style="color:#C0522A;"}: staying open and flexible, and using
  conflict to connect rather than to win.
  - *Survey item:* "Listen patiently, try to understand other people from their perspective."

**Polarity:** Violent ↔ Nonviolent.

**Measure:** Nonviolent Communication Behaviors Scale (Cheung et al., 2022). Seven items
rated 1 = Never to 4 = Frequently.

## Find Your Score {.unnumbered}

Use your password to find your score(s).

::: {.callout-note}
## What "Strength" means here
Your **Strength** is whichever of the three behaviors sits **highest against your cohort** —
your best percentile of the three, not your highest raw score. The three subscales are built
from different numbers of items and spread differently, so comparing raw means across them
would mostly describe the scales rather than you.

It is a *relative* strength, so everyone has one. It does not say you are strong on that
behavior and weak on the others, and it does not say the other two need fixing. Read it as the
place you already have something to build from.
:::

```{r}
#| label: ch10-communication-style-compute
#| include: false
comm <- dat |> transmute(Password = PASSWORD,
  `Self-connection`           = rowMeans(across(c(Self_Connect1, Self_Connect2))),
  `Authentic self-expression` = rowMeans(across(c(Self_Express1, Self_Express2))),
  `Empathic listening`        = rowMeans(across(c(Empath_Listen1, Empath_Listen2, Empath_Listen3))))

# Strength = whichever of the three sits highest against the cohort. Percentiles, not
# raw means: the three subscales have different item counts and different spreads, so
# comparing raw scores across them would mostly reflect the scales, not the mentor.
# Rounded to the values printed in the table so the label agrees with them. Ties are
# joined rather than broken - one mentor in this cohort ties on two.
pc_comm <- sapply(comm[, -1], function(x) round(dplyr::percent_rank(x) * 100))
strength <- apply(pc_comm, 1, function(r)
  paste(colnames(pc_comm)[r == max(r)], collapse = " & "))
```

```{r}
#| label: ch10-communication-style-table
#| echo: false
score_table(comm$Password, list(`Self-connection`=comm$`Self-connection`,
  `Authentic self-expression`=comm$`Authentic self-expression`,
  `Empathic listening`=comm$`Empathic listening`),
  "10. Communication Style: your NVCBS composites (1–4)",
  extra = list(Strength = strength))
```

## Locate Your Score {.unnumbered}

Locate your score(s) relative to other peer mentors!

```{r}
#| label: ch10-communication-style-plot
#| echo: false
#| fig-width: 9
#| fig-height: 3.8
norm_facets(list(`Self-connection`=comm$`Self-connection`,
  `Authentic self-expression`=comm$`Authentic self-expression`,
  `Empathic listening`=comm$`Empathic listening`), xlab="Frequency",
  scale = c(1, 4), labels = c("1\nNever","2\nRarely","3\nSometimes","4\nFrequently"))
```

## The Three Dimensions Together {.unnumbered}

Nonviolent communication is not one scale but three habits working together. This plot places
each peer mentor in the space defined by all three at once. **Drag to rotate** it and see how
self-connection, authentic self-expression, and empathic listening relate; hover a point to see
its password.

```{r}
#| label: ch10-communication-style-3d
#| echo: false
#| message: false
#| warning: false
#| fig-height: 6
suppressPackageStartupMessages(library(plotly))
set.seed(1)
jit <- function(x) x + runif(length(x), -0.05, 0.05)
comm3d <- comm |> transmute(Password,
  x = jit(`Self-connection`), y = jit(`Authentic self-expression`), z = jit(`Empathic listening`))
plot_ly(comm3d, x = ~x, y = ~y, z = ~z, text = ~Password,
  hovertemplate = "%{text}<extra></extra>",
  type = "scatter3d", mode = "markers",
  marker = list(size = 4, color = green, opacity = 0.85)) |>
  layout(scene = list(
    xaxis = list(title = "Self-connection (1-4)", range = c(1, 4)),
    yaxis = list(title = "Authentic self-expression (1-4)", range = c(1, 4)),
    zaxis = list(title = "Empathic listening (1-4)", range = c(1, 4))))
```

## Reflect and Plan

::: {.callout-warning}
## Journal
- **Score:** How does reflecting on your score help you think about who you are as a mentor?
- **Behavior:** To be the best mentor you can be, what is one new thing you will try to do this week for mentees and/or your team?
:::
